---
title: "Homework 5"
author: "Mari Sanders"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(faraway)
library(corrplot)
library(leaps)
```

a) 

```{r}

state_data <- 
  state.x77 %>% as_tibble() %>% janitor::clean_names()
state_data %>% summary()
state_data %>% 
  summarize(population_sd = sd(population, na.rm = TRUE),
            income_sd = sd(income, na.rm = TRUE),
            illiteracy_sd = sd(illiteracy, na.rm = TRUE),
            lifeexpec_sd = sd(life_exp, na.rm = TRUE),
            murder_sd = sd(murder, na.rm = TRUE),
            hsgrad_sd = sd(hs_grad, na.rm = TRUE),
            frost_sd = sd(frost, na.rm = TRUE),
            area_sd = sd(area, na.rm = TRUE))
```

b) 

```{r}
pairs(state_data)
corrplot(cor(state_data), type = "upper", diag = FALSE)

```

```{r}
par(nfrow = c(2,3))
boxplot(state_data$population, main = "Population") 
boxplot(state_data$income, main = "Income")
boxplot(state_data$illiteracy, main = "Illiteracy")
boxplot(state_data$life_exp, main = "Life Expectancy")
boxplot(state_data$murder, main = "Murder")
boxplot(state_data$hs_grad, main = "High School Grad")
boxplot(state_data$frost, main = "Frost")
boxplot(state_data$area, main = "area")
```

Life expectancy, and murder seem to have a relationship, as well as life expectancy and high school grad and life expectancy and illiteracy. There seems to be a slight relationship between life expectancy and frost. It seems like Illiteracy, population, and area
are skewed. 

c) 

```{r}
full_model <- lm(life_exp ~ ., data = state_data)
summary(full_model)

b = regsubsets(life_exp ~., data = state_data)
rs = summary(b)

plot(2:8, rs$cp, xlab = "No. of Parameters", ylab = "CP Statistic")
plot(2:8, rs$adjr2, xlab = "No. of Parameters", ylab = "Adj R2")
```

Using subsetting, the best model seems to be one that contains population, income, murder, hs_grad, and frost. 
`life_exp` = `population`$\beta_1$ + `income` $\beta_2$ + `murder` $\beta_3$ + `hs_grad` $\beta_4$ + `frost` $\beta_5$

d)

```{r}

```

