---
title: "Homework 5"
author: "Mari Sanders"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(faraway)
library(corrplot)
library(leaps)
library(glmnet)
library(patchwork)
set.seed(1)
```

a) 

```{r}

state_data <- 
  state.x77 %>% as_tibble() %>% janitor::clean_names()
state_data %>% summary() %>% knitr::kable()
state_data %>% 
  summarize(population_sd = sd(population, na.rm = TRUE),
            income_sd = sd(income, na.rm = TRUE),
            illiteracy_sd = sd(illiteracy, na.rm = TRUE),
            lifeexpec_sd = sd(life_exp, na.rm = TRUE),
            murder_sd = sd(murder, na.rm = TRUE),
            hsgrad_sd = sd(hs_grad, na.rm = TRUE),
            frost_sd = sd(frost, na.rm = TRUE),
            area_sd = sd(area, na.rm = TRUE)) %>% knitr::kable()
```

b) 

```{r}
pop <- ggplot(state_data, aes(x = log(population))) + geom_histogram()

income <- ggplot(state_data, aes(x = income)) + geom_histogram()

ggplot(state_data, aes(x = 1/(log(illiteracy) + 1))) + geom_histogram()

ggplot(state_data, aes(x = life_exp)) + geom_histogram()
 
ggplot(state_data, aes(x = hs_grad)) + geom_histogram()

ggplot(state_data, aes(x = frost)) + geom_histogram()

ggplot(state_data, aes(x = log(area))) + geom_histogram()
```


```{r}
state_data <- 
  state_data %>% 
  mutate(population = log(population), 
         illiteracy = 1/(log(illiteracy) + 1), 
         area = log(area))
```


```{r}
par(nfrow = c(2,3))

pop <- ggplot(state_data, aes(y = population)) + geom_boxplot()
income <- ggplot(state_data, aes(y = income)) + geom_boxplot() 
illiteracy <- ggplot(state_data, aes(y = illiteracy)) + geom_boxplot()  
life_exp <- ggplot(state_data, aes(y = life_exp)) + geom_boxplot()
murder <- ggplot(state_data, aes(y = murder)) + geom_boxplot()  
grad <- ggplot(state_data, aes(y = hs_grad)) + geom_boxplot()  
frost <- ggplot(state_data, aes(y = frost)) + geom_boxplot()
area <- ggplot(state_data, aes(y = area)) + geom_boxplot()

pop + income + illiteracy + life_exp + murder + grad + frost + area 
               
```

Life expectancy, and murder seem to have a relationship, as well as life expectancy and high school grad and life expectancy and illiteracy. There seems to be a slight relationship between life expectancy and frost. It seems like Illiteracy, population, and area
are skewed. 

c) 

```{r}
full_model <- lm(life_exp ~ ., data = state_data)
only_1 <- lm(life_exp ~ 1, data = state_data)
summary(full_model)



summary(step(only_1, direction = "forward", scope = formula(full_model)))

summary(step(full_model, direction = "backward"))

summary(step(full_model, direction = "both"))

x <- lm(life_exp ~ population + murder + hs_grad, data = state_data)

summary(x)
```
Doing backward, both, and regular subsetting gets the same result to include population, murder, hs_grad, and frost, area. 
`life_exp` = `population`$\beta_1$ + `murder` $\beta_2$ + `hs_grad` $\beta_3$ + `frost` $\beta_4$ 

The only "close call" variable is `frost`, but the adjusted r-squared value decreased when you take `frost` out, so I chose to keep it in the model. 

```{r}
pairs(state_data)
corrplot(cor(state_data), type = "upper", diag = FALSE)
```

There is a strong relationship between `illiteracy` and `hs_grad`. Our subset only contains `hs_grad` and not `illiteracy`. 

d) 

```{r}
data <- state_data %>% select(-life_exp)
mat <- as.matrix(data)
leaps(x = mat[,1:7], y = mat[,8], nbest = 2, method = "Cp")
```


```{r}
subsets = regsubsets(life_exp ~., data = state_data)
subset_res = summary(subsets)

subset_res
plot(subset_res$cp, xlab = "No. of Parameters", ylab = "CP Statistic")
plot(subset_res$adjr2, xlab = "No. of Parameters", ylab = "Adj R2")
```


e) 

```{r}
lambda_seq <- 10^seq(-3,0, by = 0.1)
cv_object <- cv.glmnet(as.matrix(state_data[2:8]), state_data$life_exp, lambda = lambda_seq, nfolds = 5)

cv_object
tibble(lambda = cv_object$lambda, 
       mean_cv_error = cv_object$cvm) %>% 
  ggplot(aes(x = lambda, y = mean_cv_error)) + geom_point()

cv_object$lambda.min

fit_bestcv <- glmnet(as.matrix(state_data[2:8]), state_data$life_exp, lambda = cv_object$lambda.min)
coef(fit_bestcv)

```

